---
title: "Logistic Regression Package - Group 8"
author: "Lucas Parvin, Priyadarshni Jitendrabhai Patel, & Felix Satognon "
date: "`r Sys.Date()`"
output: github_document
---
+ ChatGPT was utilized in the creation of this package: https://chat.openai.com/share/cbb8309e-df87-444a-974e-09e9f0cb91ff
+ GitHub Repository Link: https://github.com/R-4-Data-Science/FinalProject_Group8
+ **This package's website can be viewed here:** https://r-4-data-science.github.io/FinalProject_Group8/


# **Project 8 Final Project Package**

## **Accessing The Package**

1. Users need to install the "devtools" package first, if not already installed, using: `install.packages("devtools")`
2.Once "devtools" is installed, it needs to be loaded into the R environment using: `library(devtools)`
3. The final project package is hosted on GitHub. Users can install it using: `devtools::install_github("R-4-Data-Science/FinalProject_Group8")`

## Finally, load our package:

```{r}
library(package8)
```

  
## **How Our Package Works:**

*Before we start using the functions, let's generate a synthetic dataset on which we can test our functions.*

```{r}
set.seed(123)  # For reproducibility
n <- 300       # Number of observations
p <- 2         # Number of predictors

# Generating predictors
X <- matrix(rnorm(n * p), ncol = p)
colnames(X) <- c("Predictor1", "Predictor2")

# Generating response
beta_true <- c(-0.5, 0.75, 1.25) # True coefficients including intercept
logits <- cbind(1, X) %*% beta_true
prob <- 1 / (1 + exp(-logits))
y <- rbinom(n, 1, prob)
```


### **Logistic Regression Using Numerical Optimization `logistic_regression()`**

This function performs logistic regression using numerical optimization. It does not rely on existing logistic regression functions in R.This function is independent of existing logistic regression functions in R.

##### *Example Usage*
```{r}
logistic_model <- logistic_regression(X, y)
print(logistic_model)

```

### **Initial Values for Logistic Regression Optimization**
This function computes the initial values for the optimization of logistic regression coefficients using the least squares method.

##### *Example Usage*
```{r}
initial_beta <- initial_values(X, y)
print("Initial Coefficients:")
print(initial_beta)# Obtain initial coefficients
```


### **Bootstrap Confidence Intervals**
This function calculates bootstrap confidence intervals for logistic regression coefficients.

##### *Example Usage*
```{r}
ci <- bootstrap_ci(X, y)
print("Bootstrap Confidence Intervals:")
print(ci)
```

### **Plotting the Fitted Logistic Regression Curve**
This function plots the logistic regression curve based on fitted model coefficients.

##### *Example Usage*
```{r}
# Plotting the logistic curve
plot_logistic_curve(X, y, logistic_model$coefficients, predictor_index = 1)
```



### **Confusion Matrix and Metrics**
Calculate and display the confusion matrix and various metrics.


##### *Example Usage*
```{r}
fitted_probabilities <- 1 / (1 + exp(-cbind(1, X) %*% logistic_model$coefficients))
conf_matrix <- calculate_confusion_matrix(y, ifelse(fitted_probabilities > 0.5, 1, 0))
print("Confusion Matrix:")
print(conf_matrix)

metrics_list <- list(
  prevalence = calculate_prevalence(y),
  accuracy = calculate_accuracy(conf_matrix),
  sensitivity = calculate_sensitivity(conf_matrix),
  specificity = calculate_specificity(conf_matrix),
  false_discovery_rate = calculate_false_discovery_rate(conf_matrix),
  diagnostic_odds_ratio = calculate_diagnostic_odds_ratio(conf_matrix)
)

print("Metrics:")
print(metrics_list)

```


### **Plot Selected Metrics Over Various Cutoff Values**

This function plots selected metrics (e.g., accuracy, sensitivity) evaluated over a range of cutoff values. It specifies metrics of interest, like prevalence or sensitivity, and visualize them over varying cutoff values.

##### *Example Usage*
```{r}
metrics_names <- c("Prevalence", "Accuracy", "Sensitivity", "Specificity", "False Discovery Rate", "Diagnostic Odds Ratio")
for (metric in metrics_names) {
  plot_selected_metrics_over_cutoffs(fitted_probabilities, y, metric)
}

```












